---
title: "WUP_04"
author: "Michelle Benavides"
date: "July 7, 2022"
output:
  html_document:
    df_print: paged
---

```{r include=FALSE,echo=FALSE}
require(knitr)
require(tidyverse)
require(tigerstats)
file="GSScsv.csv"
mydata <- read_csv(file="GSScsv.csv")
```

# Introduction

<!--In this section you explain what you are trying to show.  Where did the data come from?  What is the research or other question you are trying to answer?!-->

The data is in the file GSScsv.csv and came from the 2018 General Social Survey.  The question to answer is to resolve this hypothesis.  
$H_0$ There is no difference between the average score on the WORDSUM test and level of schooling DEGREE  
$H_a$ There is a significant difference in average score on the WORDSUM test between at least two levels of DEGREE  

# Methods

<!--Decide on your methods:  use "variable analysis" or other appropriate descriptors.  Make sure to choose at least one graphical method and at least one numerical method.!-->

# Results

<!--Divide this section into two sub-sections:  One for your descriptive  results and one for your inferential results.!-->

## Descriptive Results

### Graphical Descriptive Results

<!--Graphical results here.  Make sure to show your code.  Provide appropriate labels for axes, giving units if possible, and provide a good title for the graph, too.  Use the graphical results to describe the patterns if any that exist in the data as focused toward the research question!-->

```{r}
basicCN <- ggplot(mydata,aes(x=DEGREE,y=WORDSUM))
```
```{r}
basicCN + geom_boxplot()
```

According to the geom boxplot, it appears as though the people with a graduate level degree did significantly better on the Wordsum test that the people in the other categories. On the other hand, the people who have a Light High School level degree had significantly worse degrees than the other people in the sample. The outliers show a range of people who have a light high school degree who scored similar to the people with a gradute degree and vice versa. 

### Numerical Descriptive Results

<!--Numerical results go here. Use the numerical results to describe the patterns if any that exist in the data as focused toward the research question!-->
```{r}
favstats(WORDSUM~DEGREE,data=mydata)
```

According to the numerical results, it seems as though the people with a Light High School degree have similar results to the results of the people that have a High School degree. While the people with a Graduate degree outranked the other people. However, the population of each degree is very different from one another. 

## Inferential Results

<!--State hypothesis clearly.  Make sure your discussion of the inferential test covers all the aspects that the test output produces, such as test statistic, p-value etc.  Make a decision about the null hypothesis, explain the assumptions on which the selected test/procedure was based, and why the chosen procedure satisfys the assumptions and is appropriate to answer the research question!-->
```{r}
model1 <- aov(WORDSUM~DEGREE,data=mydata)
summary.aov(model1)
```

The p-value 0.000236 is too small for us to think that there is no difference between any of the categories. 

```{r}
#we specify the margins starting at the bottom and proceeding clockwise - thus 8.7 is the left side margin to hold the names of the factors
par(mar = c(3, 15.7, 2, 2))
plot(TukeyHSD(model1),las=2)
```

According to these confidence intervals, the data of the High School and Graduate Degrees are significantly different, as well as Light High Schooland Graduate Degrees are also significantly different. 

# Discussion and Conclusion

<!--Discussion and conclusion here.  If you found a relationship be sure to consider whether the relationship occurs because one of the variavbles causes the other, or whether they perhasps are related for some other reason.  Watch the chapter 6 videos from the GeorgeTown videos collection.!-->

Along the process of attempting to upload data from the general survey, I became frustrated because I couldn't fully access them. However, once I did access the data I inputted a geom boxplot, numerical results, and inferential results with a TukeyHSD. The graphs and results show that once a person receives there Graduates degree they have a better chance of getting a higher score on the Wordsum test. 
